Global Climate and Atmospheric Composition of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and \Spitzer\ Phase Curve Observations

### 2018-03-30 ApJ_submit_v2/response.txt

1       x
2       x
3       x
4       x
5	x
6
7
8
9
10	x
11
12
13
14
15
16
17
18
19	x
20
21
22
23

@@

Particularly, when selecting aperture sizes, how are the pixels partially within the aperture weighted? Are they weighted by the portion of the pixel within the aperture, or is this multiplied by a sensitivity map / or the portion of an estimated PSF that falls in the pixel? This is not stated later in 3.2.2.

Some of the information of the reduction is found within Cubillos et al. 2013, which the authors have cited. However, this is mixed in with analysis specific to the object in that system. This paper also notes that the POET pipeline has several 'options' to choose from along the way in the reduction.

Therefore, I think section 2.2 should be significantly expanded (or this info should be in 3.2.2) to explain what the authors have done to reduce the Spitzer data so that it is more easily reproduced. Especially since many people (and this referee in particular) are much more familiar with PLD methods rather than BLISS mapping type methods.

We have added the following additional description to Sections 2.2 and 3.2.2 to clarify these points:

Section 2.2:
We used the POET pipeline (Stevenson et al. 2012; Cubillos et al. 2013) to reduce the Spitzer data.  The pipeline starts by identifying and flagging bad pixels using a double-iteration 4-sigma outlier rejection routine along the time axis.  This is followed by performing 2D Gaussian centroiding on each frame, which is shown to provide the most precise centers for Spitzer data \citep{Lust2014}.  The target remains centered near the sweet spot for the entire AOR in each observation, with a maximum drift of 0.1 pixels.  Next, POET uses sub-pixel (5\times-interpolated) aperture photometry \citep{Harrington2007} to subtract the background and sum the flux within a specified radius.  Chosen from a grid of apertures between 2 and 4 pixels, we find that an aperture size of 2.75 pixels minimizes the residual noise in the light curve fits.  For the background, we use an annulus with inner and outer radii of 7 and 15 pixels, respectively.  The contaminating flux from the nearby star is within the same pixel as the target, so we included it in the photometry and corrected it in the light curve fits.  A similarly strategy has been applied to successfully analyze dozens of Spitzer data sets \citep[e.g. ][]{Stevenson2010, Stevenson2012a, Stevenson2012b, Stevenson2014a, Stevenson2014b, Stevenson2016a, Stevenson2017a, Campo2011, Nymeyer2011, Cubillos2013, Blecic2013, Blecic2014, HDL2014}.

Section 3.2.2:
Warm Spitzer's primary systematic is intrapixel sensitivity variations, where the photometry depends on the precise location of the stellar center within its pixel.  We fit this systematic using the Bilinearly-Interpolated Subpixel Sensitivity (BLISS) mapping technique \citep{Stevenson2012a}.  BLISS provides a flexible, non-analytic means to effectively weight the target flux by the spatial sensitivity variations within a pixel, while simultaneously fitting for other systematics and the physical parameters of the system.  As has been demonstrated by \citet{Ingalls2016}, the POET pipeline with BLISS mapping can accurately model simulated {\em Spitzer} light curves with known physical parameters and produce reliable results.

@@

****

2. Similarly, in section 2.3 the authors state at the end of paragraph 1 "The nightly observations were corrected for bias, flat-fielding, pier-side offset, and differential atmospheric extinction". I would insist that more details are provided about this reduction. Bias and flat-fielding are straightforward and likely don't need additional explanation but the pier-side offset and the differential atmospheric extinction portion could use some more explanation:

a) Is the "pier-side offset" a singular term (constant) or is it a more general correction that's a function of the tilt of the telescope to account for flexure? Was it measured or is it calculated? etc.

Thank you for raising this point -- it was an oversight to leave the pier-side correction in the manuscript.

The C14 telescope was originally operated on a German equatorial mount, so the telescope had to flip over the RA axis in order follow a star from the east into the west.  This rotated the field on the CCD, which meant we were using a different part of the chip to make our measurements.  In case this caused a slight shift in the data not corrected by flat-fielding, we determined and applied a pier-side offset correction.  In the case of 55 Cnc, the correction was only 0.00007 mag (70 micro-mags), i.e., completely insignificant and smaller than its uncertainty.  Furthermore, last year we remounted the C14 on a horse-shoe type mount that does not require any flipping to track a star past the meridian so there is no longer any pier-side correction. We removed the mention of pier-side offset from section 2.3.

@@

The C14 telescope was originally operated on a German equatorial mount, so the telescope had to flip over the RA axis in order follow a star from the east into the west.  This rotated the field on the CCD, which meant we were using a different part of the chip to make our measurements.  In case this caused a slight shift in the data not corrected by flat-fielding, we determined and applied a pier-side offset correction.  In the case of 55 Cnc, the correction was only 0.00007 mag (70 micro-mags), i.e., completely insignificant and smaller than its uncertainty.  Furthermore, last year we remounted the C14 on a horse-shoe type mount that does not require any flipping to track a star past the meridian so there is no longer any pier-side correction. We removed the mention of pier-side offset from section 2.3.

b) I have similar questions for the differential atmospheric extinction. Is this a linear term with airmass? Quadratic? Is it measured from the data as a correction or is it calculated? What form does it take? Is it something like (X-1) * (color) where X is the airmass and color is some determination of color from catalogs?

More detail on these corrections is warranted.

We have added the following description of our correction for extinction:
    We determined seasonal-mean extinction coefficients from a linear, least-squares fits to seasonal plots of differential magnitudes versus mean air mass and computed the slope in magnitudes per air mass.

@@

We have added the following description of our correction for extinction:
    We determined seasonal-mean extinction coefficients from a linear, least-squares fits to seasonal plots of differential magnitudes versus mean air mass and computed the slope in magnitudes per air mass.

****

3. Figure 1 and associated discussion: Please include separate panels for the reference star time series. Since only 6 reference stars were used, these can all be accommodated with one extra panel with different reference stars offset from each other.

We thank the referee for this attention to detail, but we feel it would not be informative to show the reference star time series for several reasons. First, independent comparison star magnitudes are not computed in our reductions; rather, the software measures the total ADU counts in each exposure for WASP-103 and the 6 comparison stars and directly computes the [program - (comp1+comp2+comp3+comp4+comp5+comp6)/6] differential magnitudes.  Second, because the brightness variation in WASP-103 is so slight, the scatter in the program star is not significantly different than the comparison stars and so not discernible by eye from a plot.

@@

****

4. Section 3.1, paragraph 2: The authors state that they fixed the orbital period, inferior conjunction time, inclination and a/R* to the values in Southworth et al. (2015). Southworth et al. (2015) collected several ground based transits with GROND (which has an observing cadence of approximately 1 minute). However, Spitzer's observing cadence is 12 seconds for this object (and does not observe through an atmosphere), so I would think that the Spitzer values fit for these parameters would be better. The authors state that the 4.5 micron data are consistent with the Southworth et al. parameters, which justified holding the parameters fixed. My question for the authors is: Were the Spitzer parameters better? This is not mentioned, but it would not surprise me if the transit parameters were better (or at least less systematically correlated), and therefore those should be the ones used rather than the Southworth et al. parameters.

Additionally, the authors have collected transit data but all of the plots shown use y-scales that don't show the transit. I think a transit plot would be nice, just to see what the data even look like.

Southworth et al. observed a total of 18 transits of WASP-103b, so with the combined data set are able to measure a very precise transit ephemeris (their reported 1 sigma uncertainty is less than 5 seconds). We fit the Spitzer 4.5 micron transit alone and obtained a precision on the ephemeris of 30 seconds (at 1 sigma). Our precision is lower due to photon noise limitations: the star is much fainter at Spitzer observations than in the optical, and the ground-based telescopes are larger than Spitzer, and we have just one transit compared to 18 in the Southworth et al. study. For this reason, we elected to use the Southworth et al. transit parameters.

FIXME -- transit plot??

@@

****

5. Section 3.1, final paragraph: The authors state the uncertainty on the companion star flux contribution is less than 1%. 1% of what? Is the error less than 1% of the total flux in the system (since WASP-103 is bright?) Or are the authors stating that they have less than 1% photometry on the magnitude of the contaminating star. If the latter, I find that difficult to believe - getting 1% absolute photometry is hard and I would be interested in the details. If it's the former (or something else), please include a couple words to clarify.

#

We were referring to the former, and added some text to clarify this point. The sentence now reads: The uncertainty on the companion star flux contribution to the total system flux is less than 1%, which introduces negligible error in the estimated planet-to-star-flux compared to the photon noise.

@@

****

6. Figure 3: This figure doesn't seem to be very useful. The large amplitude of the systematics of Hubble observations and Spitzer observations are well known, and visualizing them here doesn't add much to the discussion. I also do not like that the panels are stacked but the x-axis is in units of "time since the first exposure", since these observations were not taken at the same time, but the visual presentation implies as such. Finally (and this is a general complaint), the unbinned Spitzer data is never shown. One of the strengths of Spitzer is it's rapid cadence, and yet Figure 3 has already binned the data to 15 minute segments. It would be useful (as has been done in other papers) to show the unbinned time series with some suitable alpha to make all the data visible and then overplot a binned version on top of that.

If the authors opt to keep Figure 3 (I think Figure 5 on its own is sufficient with some changes), I would like to see the unbinned data in a plot, and the x-axis changed to something more suitable (Time or Phase).

Several of the theorists on this paper have spoken up to say that they like the reminder of what the raw data look like, since they don't work with the data directly. We therefore decided to keep the figure, but made the suggested changes.

### 2018-03-30 ApJ_submit_v2/ms.tex
ref response

\documentclass[twocolumn]{aastex61}

@@

The data were reduced with the POET pipeline \citep{stevenson12, cubillos13}. We performed aperture photometry with an aperture size of 2.75 pixels (chosen from a grid of apertures between 2 and 4 pixels to minimize the residual noise in the light curve fits). We masked pixels that were flagged in the bad pixel mask provided in the ancillary data for the observations. The target centroid was determined with a two-dimensional Gaussian fit.  The target remains centered near the sweet spot for the entire AOR in each observation, with a maximum drift of 0.1 pixels. We estimated and subtracted the background from an annulus with a radius of 7 to 15 pixels from the centroid. The contaminating flux from the nearby star is within the same pixel as the target, so we included it in the photometry and corrected it in the light curve fits.

\subsection{Photometric Monitoring}
To assess how stellar activity might impact the phase curve observations, we monitored WASP-103's photometric variability over 158 nights during 2014 - 2016 with the Tennessee State University Celestron 14-inch (C14) automated imaging telescope (AIT), located at Fairborn Observatory in southern Arizona \citep[][]{henry99}.  The observations of WASP-103 were made in the Cousins R passband with an SBIG STL-1001E CCD camera.  Each observation consisted of 4--10 consecutive exposures on WASP-103 along with several dozen comparison stars in the same field. The individual consecutive frames were co-added and reduced to differential magnitudes (i.e., WASP-103 minus the mean brightness of the six best comparison stars). The nightly observations were corrected for bias, flat-fielding, pier-side offset, and differential atmospheric extinction.

The photometric analyses are summarized for each observing season in Table\,\ref{tab:photometry}.  The standard deviation of WASP-103's brightness in each season is given in column~4.  The mean of the three standard deviations is 0.0058~mag, which is comparable to the mean standard deviation of the six comparison stars (0.0055~mag).  To maximize the possibility of detecting WASP-103's rotation, we normalized the photometry such that each observing season has the same mean, thereby removing any long-term variability in WASP-103 and/or the comparison stars (Figure\,\ref{fig:photometry}, top panel).  To estimate the stellar rotation period, we performed a periodogram analysis of the normalized data set based on least-squares fitting of sine curves.  The resulting frequency spectrum and the phase curve computed with the best period are shown in the middle and lower panels of Figure\,\ref{fig:photometry} respectively. The best-fit period is 6.814 days, which agrees closely with the estimated stellar rotation period of 6.855 days \citep[based on the projected stellar rotation velocity reported in][]{gillon14}. The peak-to-peak variability amplitude is 0.005 mag.  Based on the formalism in \cite{zellem17}, we calculate that this variability will bias the measured eclipse depth by $\lesssim10$\,parts per million (ppm) from epoch to epoch, which is well below the photon-limited precision of our measurements.

@@

\subsubsection{\Spitzer\ Systematics}
We fit the \Spitzer\ systematics with the POET pipeline, which uses the BLISS mapping technique \citep{stevenson12}. This approach creates a map of the intrapixel sensitivity while simultaneously fitting for other systematics and the physical parameters of the system. The sensitivity map is determined by bilinear interpolation over a grid of knots centered on the stellar flux. Each knot's sensitivity is calculated from the residuals to the light curve fit: the higher the flux values for data points near a given knot, the higher the detector sensitivity is at that position.  To avoid overfitting, we chose the grid scale such that bilinear interpolation performed better than nearest neighbor interpolation. For the $3.6\,\mu$m data, the grid scale was 0.008 pixel in both $x$ and $y$. For $4.5\,\mu$m, the scale was 0.022 pixel.  In addition to the intrapixel sensitivity variation, we fit the data for a linear trend in time. We tested a quadratic trend but did not find significant evidence for the additional model complexity based on the BIC.

>>>>

\documentclass[twocolumn, trackchanges]{aastex61}

@@

The data were reduced with the POET pipeline \citep{stevenson12, cubillos13}.  \added{The pipeline starts by identifying and flagging bad pixels using a double-iteration 4-sigma outlier rejection routine along the time axis.  This is followed by performing 2D Gaussian centroiding on each frame, which is shown to provide the most precise centers for Spitzer data \citep{Lust2014}.  The target remains centered near the sweet spot for the entire AOR in each observation, with a maximum drift of 0.1 pixels.  Next, POET uses sub-pixel (5$\times$-interpolated) aperture photometry \citep{Harrington2007} to subtract the background and sum the flux within a specified radius.  Chosen from a grid of apertures between 2 and 4 pixels, we find that an aperture size of 2.75 pixels minimizes the residual noise in the light curve fits.  For the background, we use an annulus with inner and outer radii of 7 and 15 pixels, respectively.  The contaminating flux from the nearby star is within the same pixel as the target, so we included it in the photometry and corrected it in the light curve fits.  A similarly strategy has been applied to successfully analyze dozens of Spitzer data sets \citep[e.g. ][]{Stevenson2010, Stevenson2012a, Stevenson2012b, Stevenson2014a, Stevenson2014b, Stevenson2016a, Stevenson2017a, Campo2011, Nymeyer2011, cubillos13, Blecic2013, Blecic2014, HDL2014}.}

%We performed aperture photometry with an aperture size of 2.75 pixels (chosen from a grid of apertures between 2 and 4 pixels to minimize the residual noise in the light curve fits). We masked pixels that were flagged in the bad pixel mask provided in the ancillary data for the observations. The target centroid was determined with a two-dimensional Gaussian fit.  The target remains centered near the sweet spot for the entire AOR in each observation, with a maximum drift of 0.1 pixels. We estimated and subtracted the background from an annulus with a radius of 7 to 15 pixels from the centroid. The contaminating flux from the nearby star is within the same pixel as the target, so we included it in the photometry and corrected it in the light curve fits.

\subsection{Photometric Monitoring}
To assess how stellar activity might impact the phase curve observations, we monitored WASP-103's photometric variability over 158 nights during 2014 - 2016 with the Tennessee State University Celestron 14-inch (C14) automated imaging telescope (AIT), located at Fairborn Observatory in southern Arizona \citep[][]{henry99}.  The observations of WASP-103 were made in the Cousins R passband with an SBIG STL-1001E CCD camera.  Each observation consisted of 4--10 consecutive exposures on WASP-103 along with several dozen comparison stars in the same field. The individual consecutive frames were co-added and reduced to differential magnitudes (i.e., WASP-103 minus the mean brightness of the six best comparison stars). The nightly observations were corrected for bias, flat-fielding, \deleted{pier-side offset,} and differential atmospheric extinction.  \added{We determined seasonal-mean extinction coefficients from a linear, least-squares fits to seasonal plots of differential magnitudes versus mean air mass and computed the slope in magnitudes per air mass.}

The photometric analyses are summarized for each observing season in Table\,\ref{tab:photometry}.  The standard deviations of a single observation with respect to the corresponding seasonal means are given in column~4; \added{the mean of the three standard deviations is 0.0058~mag, suggesting there is little night-to-night variation in WASP-103.  The three seasonal mean brightness values given in column 5 scatter about their grand mean with a standard deviation of 0.0036 mag, but we note that the most discrepant mean is from the third season, for which we have only partial coverage. Therefore, our results do not completely rule out low-level, year-to-year variability of $<0.001$~mag.}

%The standard deviation of WASP-103's brightness in each season is given in column~4.  The mean of the three standard deviations is 0.0058~mag, which is comparable to the mean standard deviation of the six comparison stars (0.0055~mag).

To maximize the possibility of detecting WASP-103's rotation, we normalized the photometry such that each observing season has the same mean, thereby removing any long-term variability in WASP-103 and/or the comparison stars (Figure\,\ref{fig:photometry}, top panel).  To estimate the stellar rotation period, we performed a periodogram analysis of the normalized data set based on least-squares fitting of sine curves.  The resulting frequency spectrum and the phase curve computed with the best period are shown in the middle and lower panels of Figure\,\ref{fig:photometry} respectively. The best-fit period is 6.814 days, which agrees closely with the estimated stellar rotation period of 6.855 days \citep[based on the projected stellar rotation velocity and stellar radius reported in][]{gillon14}. The peak-to-peak variability amplitude is 0.005 mag.  Based on the formalism in \cite{zellem17}, we calculate that this variability will bias the measured eclipse depth by $\lesssim10$\,parts per million (ppm) from epoch to epoch, which is well below the photon-limited precision of our measurements.

@@

\subsubsection{\Spitzer\ Systematics}
\added{Warm Spitzer's primary systematic is intrapixel sensitivity variations, where the photometry depends on the precise location of the stellar center within its pixel.  We fit this systematic using the Bilinearly-Interpolated Subpixel Sensitivity (BLISS) mapping technique \citep{stevenson12}.  BLISS provides a flexible, non-analytic means to effectively weight the target flux by the spatial sensitivity variations within a pixel, while simultaneously fitting for other systematics and the physical parameters of the system. As demonstrated by \citet{Ingalls2016}, the POET pipeline with BLISS mapping can accurately model simulated \Spitzer\ light curves with known physical parameters and produce reliable results.}

The BLISS sensitivity map is determined by bilinear interpolation over a grid of knots centered on the stellar flux. Each knot's sensitivity is calculated from the residuals to the light curve fit: the higher the flux values for data points near a given knot, the higher the detector sensitivity is at that position.  To avoid overfitting, we chose the grid scale such that bilinear interpolation performed better than nearest neighbor interpolation. For the $3.6\,\mu$m data, the grid scale was 0.008 pixel in both $x$ and $y$. For $4.5\,\mu$m, the scale was 0.022 pixel.  In addition to the intrapixel sensitivity variation, we fit the data for a linear trend in time. We tested a quadratic trend but did not find significant evidence for the additional model complexity based on the BIC.
